---
title: "ReproducibleResearchproject2"
author: "Saydaliev"
date: "4/1/2020"
output: html_document
---
```{r}
library("data.table")
library("ggplot2")
```


# Import and Convert Data 
```{r}
setwd("~/Documents/DataScienceCoursera/Reproducible Research")
stormDF <- read.csv("repdata-data-StormData.csv.bz2")
stormDT <- as.data.table(stormDF)
```


# Column Names 
```{r}
colnames(stormDT)
```


#Data Subset 
```{r}
results="hide"
```


#Remove Columns 

```{r}
cols2Remove <- colnames(stormDT[, !c("EVTYPE"
                                     , "FATALITIES"
                                     , "INJURIES"
                                     , "PROPDMG"
                                     , "PROPDMGEXP"
                                     , "CROPDMG"
                                     , "CROPDMGEXP")])

```


#Removing Columns 

```{r}
stormDT[, c(cols2Remove) := NULL]
```


# Fatalities or Injured 

```{r}
stormDT <- stormDT[(EVTYPE != "?" & 
                      (INJURIES > 0 | FATALITIES > 0 | PROPDMG > 0 | CROPDMG > 0)), c("EVTYPE"
                                                                                      , "FATALITIES"
                                                                                      , "INJURIES"
                                                                                      , "PROPDMG"
                                                                                      , "PROPDMGEXP"
                                                                                      , "CROPDMG"
                                                                                      , "CROPDMGEXP") ]
```



#Making the PROPDMGEXP and CROPDMGEXP columns cleaner so they can be used to calculate property and crop cost.

```{r}
results="hide"

```



# Change all damage exponents to uppercase.

```{r}
cols <- c("PROPDMGEXP", "CROPDMGEXP")
stormDT[,  (cols) := c(lapply(.SD, toupper)), .SDcols = cols]
```



# Map property damage alphanumeric exponents to numeric values.

```{r}
propDmgKey <-  c("\"\"" = 10^0,
                 "-" = 10^0, 
                 "+" = 10^0,
                 "0" = 10^0,
                 "1" = 10^1,
                 "2" = 10^2,
                 "3" = 10^3,
                 "4" = 10^4,
                 "5" = 10^5,
                 "6" = 10^6,
                 "7" = 10^7,
                 "8" = 10^8,
                 "9" = 10^9,
                 "H" = 10^2,
                 "K" = 10^3,
                 "M" = 10^6,
                 "B" = 10^9)
```



# Map crop damage alphanumeric exponents to numeric values

```{r}
cropDmgKey <-  c("\"\"" = 10^0,
                 "?" = 10^0, 
                 "0" = 10^0,
                 "K" = 10^3,
                 "M" = 10^6,
                 "B" = 10^9)
stormDT[, PROPDMGEXP := propDmgKey[as.character(stormDT[,PROPDMGEXP])]]
stormDT[is.na(PROPDMGEXP), PROPDMGEXP := 10^0 ]
stormDT[, CROPDMGEXP := cropDmgKey[as.character(stormDT[,CROPDMGEXP])] ]
stormDT[is.na(CROPDMGEXP), CROPDMGEXP := 10^0 ]

```



###Making Economic Cost Columns

```{r}
stormDT <- stormDT[, .(EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, propCost = PROPDMG * PROPDMGEXP, CROPDMG, CROPDMGEXP, cropCost = CROPDMG * CROPDMGEXP)]
```



###Calcuating Total Property and Crop Cost

```{r}
totalCostDT <- stormDT[, .(propCost = sum(propCost), cropCost = sum(cropCost), Total_Cost = sum(propCost) + sum(cropCost)), by = .(EVTYPE)]
totalCostDT <- totalCostDT[order(-Total_Cost), ]
totalCostDT <- totalCostDT[1:10, ]
head(totalCostDT, 5)
```



###Calcuating Total Fatalities and Injuries

```{r}
totalInjuriesDT <- stormDT[, .(FATALITIES = sum(FATALITIES), INJURIES = sum(INJURIES), totals = sum(FATALITIES) + sum(INJURIES)), by = .(EVTYPE)]
totalInjuriesDT <- totalInjuriesDT[order(-FATALITIES), ]
totalInjuriesDT <- totalInjuriesDT[1:10, ]
head(totalInjuriesDT, 5)

```


##Results

###Events that are Most Harmful to Population Health

```{r}
bad_stuff <- melt(totalInjuriesDT, id.vars="EVTYPE", variable.name = "bad_thing")
head(bad_stuff, 5)
```


###Create chart
```{r}
healthChart <- ggplot(bad_stuff, aes(x=reorder(EVTYPE, -value), y=value))
```


###Plot data as bar chart

```{r}
healthChart = healthChart + geom_bar(stat="identity", aes(fill=bad_thing), position="dodge")
```


# Format y-axis scale and set y-axis label
```{r}
healthChart = healthChart + ylab("Frequency Count") 
```


# Set x-axis label
```{r}
healthChart = healthChart + xlab("Event Type") 
```


# Rotate x-axis tick labels 
```{r}
healthChart = healthChart + theme(axis.text.x = element_text(angle=45, hjust=1))
```


# Set chart title and center it

```{r}
healthChart = healthChart + ggtitle("Top 10 US Killers") + theme(plot.title = element_text(hjust = 0.5))
healthChart
```


###Events that have the Greatest Economic Consequences

```{r}
econ_consequences <- melt(totalCostDT, id.vars="EVTYPE", variable.name = "Damage_Type")
head(econ_consequences, 5)
```



# Create chart
```{r}
econChart <- ggplot(econ_consequences, aes(x=reorder(EVTYPE, -value), y=value))
```


# Plot data as bar chart
```{r}
econChart = econChart + geom_bar(stat="identity", aes(fill=Damage_Type), position="dodge")
```

# Format y-axis scale and set y-axis label
```{r}
econChart = econChart + ylab("Cost (dollars)") 
```

# Set x-axis label
```{r}
econChart = econChart + xlab("Event Type") 
```

# Rotate x-axis tick labels 
```{r}
econChart = econChart + theme(axis.text.x = element_text(angle=45, hjust=1))
```

# Set chart title and center it

```{r}
econChart = econChart + ggtitle("Top 10 US Storm Events causing Economic Consequences") + theme(plot.title = element_text(hjust = 0.5))
econChart
```






